Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-12-26 manuscript/manuscript.tex
Update conclusions

We have developed a new methodology to compute the gravitational fields
generated by a tesseroid with a density given by a continuous function of depth.
It numerically solves the integrals that define the gravitational potential,
its gradient, and the Marussi tensor components through the Gauss-Legendre
Quadrature (GLQ).
The accuracy of the numerical integration is automatically controlled by an adaptive
discretization algorithm and a new density-based discretization algorithm.
The former divides the tesseroid in half if the ratio of the distance to the computation
point and the size of the tesseroid is lower than a predefined distance-size ratio $D$.
This algorithm minimises the integration error when the computation point is close to
the tesseroid.
Nevertheless, the adaptive discretization alone is not sufficient to guarantee the
accuracy of the method in case of tesseroids with variable density.

@@

Because there is no direct relation between the values assigned to $D$ and $\delta$ and
the error of the computation, we empirically determined the optimum values for these
parameters.
These values minimize the computational load while maintaining the numerical error below
0.1\% of an analytical solution.
The density functions used to establish the optimum values were a linear and an
exponential function.
The linear function represents the smoothest variation of density and does not require
density-based discretization at all.
We analysed the error for exponential functions ranging from smooth to sharp variations
in depth to test the accuracy of the density-based discretization.
Through this analysis, we have obtained optimal values for the distance-size ratio of
1, 2 and 8 for the potential, its gradient, and the Marussi tensor components,
respectively.
These values are in agreement with previous results for a homogeneous density tesseroid
\citep{Uieda2016}.
The optimum $\delta$ values are 0.2 for the potential and its gradient components, while
a $\delta = 0.01$ must be used for the Marussi tensor components in order to ensure a
0.1\% precision.
A wide range of exponential functions were used to infer these results and they are
conservative estimates, chosen to prioritize accuracy over speed of computations.
Therefore, they can be safely generalized to other common depth-density relations, like
quadratic functions, without loss of accuracy.

>>>>

We have developed a new methodology to compute the gravitational fields
generated by a tesseroid with a density given by a continuous function of depth.
It numerically solves the integrals that define the gravitational potential and
its gradient through the Gauss-Legendre Quadrature (GLQ).
The accuracy of the numerical integration is automatically controlled by a two
dimensional adaptive discretization algorithm and a new density-based discretization
algorithm.
The former divides the tesseroid in half if the ratio of the distance to the computation
point and the latitudinal or the longitudinal size of the tesseroid is lower than
a predefined distance-size ratio $D$.
This algorithm minimises the integration error when the computation point is close to
the tesseroid.
Nevertheless, the adaptive discretization alone is not sufficient to guarantee the
accuracy of the method in case of tesseroids with variable density.

@@

Because there is no direct relation between the values assigned to $D$ and $\delta$ and
the error of the computation, we empirically determined the optimum values for these
parameters.
These values minimize the computational load while maintaining the numerical error below
0.1\% of an analytical solution.
The density functions used to establish the optimum values were a linear, an exponential
and a sinusoidal function.
The linear function represents the smoothest variation of density and does not require
density-based discretization at all.
We analysed the error for exponential functions ranging from smooth to sharp variations
in depth to test the accuracy of the density-based discretization.
And the error for sinusoidal functions with different wavelengths.
Through this analysis, we have obtained optimal values for the distance-size ratio of
1 and 2.5 for the potential an its gradient, respectively.
While values of $\delta$ equal to 0.1 and 0.01 are needed in order to guarantee a 0.1\%
accuracy for an exponential and a sinusoidal density, respectively.
These results could be extrapolated: a $\delta$ value equal to 0.01 is needed in order
to achieve the accuracy threshold in case of an arbitrary continuous density function.
Although we encourage to perform similar accuracy tests for functions other than the
ones we studied, this results could be safely generalized to other common depth-density
relations among geophysical applications, like quadratic functions, without loss of
accuracy.
